---
title: "FED_UNCERTAINTY"
author: "Luis Luarte"
date: "4/18/2022"
output: html_document
---

# Load libs

```{r}
pacman::p_load(
  tidyverse,
  cowplot,
  lme4,
  lmerTest,
  ggrepel,
  ggplot2
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data

```{r, echo = FALSE}
# intake
csv_list <- list.files(
  path = "../raw_data",
  recursive = TRUE,
  pattern = "\\.CSV$",
  full.names = TRUE
)
data_raw <- csv_list %>%
  set_names() %>%
  map_dfr(
    read_csv,
    .id = "file_name"
  ) %>%
  filter(animal != 111) # test animal code

# weight
weight_path <- list.files(
  path = "../raw_data/weights",
  recursive = FALSE,
  pattern = ".csv",
  full.names = TRUE
)
weight_data <- weight_path %>% 
  set_names() %>%
  map_dfr(
    read_csv,
    .id = "file_name"
  )
weight_data <- weight_data %>% 
  select(-file_name) %>% 
  mutate(
    date = lubridate::ymd(date)
  ) %>%
  gather(
    .,
    "animal",
    "weight",
    -date
  )
```

# Data pre-proc

```{r}
df <- data_raw %>%
  select(-numJamClears) %>% # useless col
  mutate(
    complete_date = as.POSIXct(time, origin = "1970-01-01"),
    date = as.Date(as.POSIXct(time, origin = "1970-01-01")),
    hour = hms::as_hms(lubridate::ymd_hms(complete_date)),
    isolated_hour = lubridate::hour(complete_date),
    animal = as.factor(animal),
    pellets = as.numeric(pellets),
    protocol = as.factor(protocol)
  ) %>%
  filter(date >= "2022-01-01")

# set experimental condition
df <- df %>%
  mutate(
    protocol = if_else(
      date <= "2022-03-28",
      "baseline",
      if_else(
        animal %in% c(321, 322, 324, 327),
        "control",
        "experimental"
      )
    )
  )
weight_data <- weight_data %>% 
    mutate(
    protocol = if_else(
      date <= "2022-03-28",
      "baseline",
      if_else(
        animal %in% c(321, 322, 324, 327),
        "control",
        "experimental"
      )
    )
  )

# remove test pellets
# pellet counter reset every 24 hours
df <- df %>%
  group_by(animal, date) %>%
  mutate(pellets = seq(1, n(), 1))

# save
saveRDS(data_raw, "df_raw.rds")
saveRDS(df, "df.rds")
```

# Start data analysis

```{r}
# load data
df <- readRDS("df.rds")
```

# Count pellets consumed every day

```{r}
pellets_consumed <- df %>%
  group_by(animal, date, protocol) %>%
  summarise(pellets_consumed = n()) %>%
  ungroup()
```

# Compute baseline intake

```{r}
baseline <- pellets_consumed %>%
  filter(
    protocol == "baseline",
    date <= "2022-03-28",
    date >= "2022-03-27"
  ) %>% 
  group_by(animal) %>% 
  summarise(
    baseline_intake = mean(pellets_consumed),
    baseline_err = sd(pellets_consumed) / sqrt(n())
  )
```

# Compute baseline weight

```{r}
weight_data_baseline <- weight_data %>%
  filter(date == "2022-03-18") %>% 
  group_by(protocol, animal) %>% 
  summarise(baseline_weight = mean(weight)) %>%
  ungroup() %>%
  select(-protocol)

merge_weight <- weight_data_baseline %>% 
  left_join(weight_data, by = c("animal")) %>% 
  filter(protocol != "baseline") %>%
  mutate(
    perc_diff = (weight - baseline_weight) / ((weight + baseline_weight) / 2) * 100
  )
```

# Compute experimental period intake

```{r}
experimental <- pellets_consumed %>% 
  filter(
    protocol != "baseline"
  )
```

# Merge experimental and baseline intake data

```{r}
merged_data <- baseline %>% 
  left_join(experimental)
```
# Compute intake in dark/light periods

```{r}
n_days_exp <- df %>% filter(protocol != "baseline")
n_days_exp <- length(unique(n_days_exp$date))
circadian_intake <- df %>% 
  filter(protocol != "baseline") %>% 
  mutate(
    day_night = as.factor(
      if_else(isolated_hour >= 12 && isolated_hour <= 23,
              "Dark", "Light")
    )
  ) %>% 
  group_by(animal, day_night, protocol) %>% 
  summarise(pellets_consumed = n() / n_days_exp)
```

# [1.0] plot: intake per experimental group

```{r}
merged_data %>% 
  filter(date >= "2022-03-30", date < "2022-04-18") %>%
  ggplot(aes(
    date,
    pellets_consumed,
    color = protocol
  )) +
  geom_line(color = "gray70") +
  geom_point() +
  geom_line(
    inherit.aes = FALSE,
    aes(
      date,
      baseline_intake
    ),
    color = "gray50"
  ) +
  facet_wrap(~animal) +
  xlab("Date") +
  ylab("Pellets consumed per day") +
  theme(text = element_text(size = 20)) +
  labs(color = "Control/Experimental") +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_bw() +
  theme(strip.background = element_rect(colour="black",
                                        fill="white"))
```
# [1.1] plot: intake per experimental group means

```{r}
# individual means
animal_intake <- merged_data %>% 
  group_by(animal, protocol) %>% 
  summarise(mean_intake = mean(pellets_consumed)) %>% 
  ungroup()
# group means
group_intake <- animal_intake %>% 
  group_by(protocol) %>% 
  summarise(
    mean_group_intake = mean(mean_intake)
    ) %>% 
  mutate(
    err = sd(mean_group_intake) / sqrt(n())
  ) %>% 
  ungroup() 
#plot
group_intake %>% 
  ggplot(aes(
    protocol,
    mean_group_intake,
    ymin = mean_group_intake - err,
    ymax = mean_group_intake + err,
    fill = protocol
  )) + 
  geom_col(width = 0.5) +
  geom_errorbar(width = 0.3) +
  geom_point(
    inherit.aes = FALSE,
    aes(protocol, mean_intake),
    data = animal_intake
  ) +
  xlab("") +
  ylab("Mean pellets consumed per day") +
  theme(text = element_text(size = 20)) +
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  scale_x_discrete(labels = c("control" = "Control", "experimental" = "Experimental")) + 
  theme_bw() +
  theme(strip.background = element_rect(colour="black",
                                        fill="white"),
        legend.position = "none")
```


# [2.0] plot: weights per animal

```{r}
merge_weight %>% 
  ggplot(
    aes(
      date,
      weight,
      color = protocol
    )
  ) +
  geom_line(color = "gray70") +
  geom_point() +
  geom_line(
    inherit.aes = FALSE,
    aes(
      date,
      baseline_weight
    ),
    color = "gray50"
  ) +
  facet_wrap(~animal) +
  xlab("Date") +
  ylab("Weight") +
  theme(text = element_text(size = 20)) +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_bw() +
  theme(strip.background = element_rect(colour="black",
                                        fill="white"))
```

# [2.1] plot: weight per group

```{r}
# individual means
animal_weight <- merge_weight %>% 
  group_by(animal, protocol) %>% 
  summarise(mean_weight = mean(perc_diff)) %>% 
  ungroup()
# group means
group_weight <- animal_weight %>% 
  group_by(protocol) %>% 
  summarise(
    mean_group_weight = mean(mean_weight)
    ) %>% 
  mutate(
    err = sd(mean_group_weight) / sqrt(n())
  ) %>% 
  ungroup() 
# plot
group_weight %>% 
  ggplot(aes(
    protocol,
    mean_group_weight,
    ymin = mean_group_weight - err,
    ymax = mean_group_weight + err,
    fill = protocol
  )) + 
  geom_col(width = 0.5) +
  geom_errorbar(width = 0.3) +
  geom_point(
    inherit.aes = FALSE,
    aes(protocol, mean_weight),
    data = animal_weight
  ) +
  xlab("") +
  ylab("Mean percent change in weight") +
  theme(text = element_text(size = 20)) +
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  scale_x_discrete(labels = c("control" = "Control", "experimental" = "Experimental")) + 
  theme_bw() +
  theme(strip.background = element_rect(colour="black",
                                        fill="white"),
        legend.position = "none")
```

# [3.0] plot: intake in dark/light periods
```{r}
circadian_intake_group <- circadian_intake %>% 
  group_by(protocol, day_night) %>% 
  summarise(
    mean_intake = mean(pellets_consumed)
    ) %>% 
  ungroup() %>% 
  mutate(
    err = sd(mean_intake) / sqrt(n())
  )
# plot
circadian_intake_group %>% 
  ggplot(aes(
    day_night,
    mean_intake,
    ymin = mean_intake - err,
    ymax = mean_intake + err,
    fill = day_night
  )) +
  geom_col() +
  geom_errorbar(width = 0.3) +
  geom_point(
    inherit.aes = FALSE,
    aes(
      day_night,
      pellets_consumed
    ),
    data = circadian_intake
  ) +
  facet_wrap(~protocol) +
  xlab("") +
  ylab("Total pellets consumed") +
  theme(text = element_text(size = 20)) +
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  scale_x_discrete(labels = c("control" = "Control", "experimental" = "Experimental")) + 
  theme_bw() +
  theme(strip.background = element_rect(colour="black",
                                        fill="white"),
        legend.position = "none")

```
