---
title: "thesis_data_analysis"
author: "Luis Luarte"
date: '2022-06-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load libs

```{r}
pacman::p_load(
  tidyverse,
  ggplot2,
  rstudioapi,
  devtools
)

# source lickometer library
devtools::source_url("https://github.com/lab-cpl/lickometer-library/blob/main/src/lickometer_functions_compilate.R?raw=TRUE")

```

# load data sets

This data sets correspond to experiments where FED food delivery was random, and
motivation was tested using a progressive ratio in lickometer

```{r}
# set working directory where this file is located
cur_dir <- dirname(getSourceEditorContext()$path)
setwd(cur_dir)

# lickometer data for both pool with low and high uncertainty
lickometer_uncertainty <- readRDS("../raw_data/lickometer_data.rds")

# fed data for both pool with low and high uncertainty
fed_high_uncertainty <- readRDS("../raw_data/high_uncertainty.rds")
fed_low_uncertainty <- readRDS("../raw_data/low_uncertainty.rds")

# merge both fed data sets
fed_uncertainty <- bind_rows(
  fed_high_uncertainty,
  fed_low_uncertainty
)

# weight data
weight_uncertainty <- readRDS("../raw_data/weights/weights.rds")
```

# data processing

## intake

```{r}
# this data proc is to plot intake

# first get intake over days per mouse
intake_over_days <- fed_uncertainty %>% 
  # add light/dark periods
  mutate(
    light_dark = case_when(
      hour_of_day >= 11 & hour_of_day <= 23 ~ "dark",
      TRUE ~ "light"
    )
  ) %>% 
  group_by(
    animal,
    ymd,
    experiment_id,
    protocol
  ) %>% 
  summarise(
    pellets_per_day = max(pellets)
  ) %>% 
  ungroup() %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    pellets_cumm = cumsum(pellets_per_day)
  ) %>% 
  filter(
    protocol %in% c("experimental", "control")
  ) %>% 
  ungroup() %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    session = as.numeric(as.factor(ymd))
  ) %>% 
  # max number of sessions for all pools
  # 246 was euthanize
  filter(session <= 35, animal != 246) %>% 
  # group both controls together
  mutate(
    group = case_when(
      protocol == "control" ~ "control",
      protocol == "experimental" &
        experiment_id == "high_uncertainty_fed" ~ "experimental_high_unc",
      protocol == "experimental" &
        experiment_id == "low_uncertainty_fed" ~ "experimental_low_unc"
    )
  )

# cummulative intake over days summarized to present data based on
# experimental protocol
intake_over_days_group <- intake_over_days %>% 
  group_by(
    group,
    session
  ) %>% 
  summarise(
    pellets_cumm_group = mean(pellets_cumm),
    err = sd(pellets_cumm) / sqrt(n())
  )

# mean intake per group
intake_means <- intake_over_days %>% 
  group_by(
    group,
    session,
    animal
  ) %>% 
  summarise(
    pellets_mean = mean(pellets_per_day)
  ) %>% 
  ungroup() %>% 
  group_by(
    group,
    session
  ) %>% 
  summarise(
    pellets_mean_session = mean(pellets_mean)
  ) %>% 
  ungroup() %>% 
  group_by(
    group
  ) %>% 
  summarise(
    pellets_mean_group = mean(pellets_mean_session),
    err = sd(pellets_mean_session) / sqrt(n())
  )

# intake with light dark periods
intake_over_days_circ <- fed_uncertainty %>% 
  # add light/dark periods
  mutate(
    light_dark = case_when(
      hour_of_day >= 11 & hour_of_day <= 23 ~ "dark",
      TRUE ~ "light"
    )
  ) %>% 
  group_by(
    animal,
    ymd,
    experiment_id,
    protocol,
    light_dark
  ) %>% 
  summarise(
    pellets_per_day = max(pellets)
  ) %>% 
  ungroup() %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    pellets_cumm = cumsum(pellets_per_day)
  ) %>% 
  filter(
    protocol %in% c("experimental", "control")
  ) %>% 
  ungroup() %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    session = as.numeric(as.factor(ymd))
  ) %>% 
  # max number of sessions for all pools
  # 246 was euthanize
  filter(session <= 35, animal != 246) %>% 
  # group both controls together
  mutate(
    group = case_when(
      protocol == "control" ~ "control",
      protocol == "experimental" &
        experiment_id == "high_uncertainty_fed" ~ "experimental_high_unc",
      protocol == "experimental" &
        experiment_id == "low_uncertainty_fed" ~ "experimental_low_unc"
    )
  )

# inter-pellet recovery time
inter_pellet_recovery <- fed_uncertainty %>% 
  mutate(
    session = as.numeric(as.factor(ymd))
  ) %>% 
  group_by(animal) %>% 
  mutate(
    inter_pellet_recovery = (hms - lag(hms)) - delay
  ) %>% 
  # na here are the first pellet which is set while leaving the FED inside the cage
  drop_na() %>% 
  ungroup() %>% 
  # remove double pellets > 3% chance
  # double pellets generate events that are faster than delay
  filter(
    inter_pellet_recovery > 0
  )

# now we want inter-pellet recovery estimate by animal and session
inter_pellet_recovery_ind <- inter_pellet_recovery %>% 
  group_by(
    animal,
    session,
    protocol,
    experiment_id
  ) %>% 
  summarise(
    mean_recovery = mean(inter_pellet_recovery)
  ) %>% 
  # now session is collapsed to obtain an estimate per animal
  ungroup() %>% 
  group_by(
    animal,
    protocol,
    experiment_id
  ) %>% 
  summarise(
    mean_recovery_id = mean(mean_recovery)
  )
# inter-pellets recovery by group
inter_pellet_recovery_group <- inter_pellet_recovery_ind %>% 
  ungroup() %>% 
  group_by(
    protocol,
    experiment_id
  ) %>% 
  summarise(
    mean_recovery_group = mean(mean_recovery_id),
    err = sd(mean_recovery_id) / sqrt(n())
  )
  
```

## weight 
```{r}
weight_uncertainty <- readRDS("../raw_data/weights/weights.rds")
# create control and experimental groups
weight_uncertainty <- weight_uncertainty %>% 
  mutate(
    group = case_when(
      animal %in% c(320, 323, 325, 326) ~ "experimental_high_unc",
      animal %in% c(234, 235, 236, 245, 265) ~ "experimental_low_unc",
      TRUE ~ "control"
    )
  )

# add groups
# first we get them from fed_uncerainty data
groups <- fed_uncertainty %>% 
  select(
    animal,
    protocol,
    ymd
  ) %>% 
  distinct(., ymd, .keep_all = TRUE)
# add the group into weight data
weight_uncertainty <- weight_uncertainty %>% 
  left_join(groups, by = c("animal", "date" = "ymd")) %>% 
  filter(protocol %in% c("experimental", "control"))

# calculate percent difference between days
weight_uncertainty <- weight_uncertainty %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    init_weight = weight[1],
    delta_bw = ((weight - init_weight) / ((weight + init_weight) / 2)) * 100,
    cumm_delta_bw = cumsum(delta_bw),
    session = as.numeric(as.factor(date))
  ) %>% 
  # right number of weight sessions
  filter(session <= 16)

# get means per group per session
weight_uncertainty_group <- weight_uncertainty %>% 
  group_by(
    session,
    group
  ) %>% 
  summarise(
    weight_mean = mean(delta_bw),
    err = sd(weight) / sqrt(n())
  )

# get mean per group
weight_uncertainty_group_ <- weight_uncertainty_group %>% 
  group_by(
    group
  ) %>% 
  summarise(
    weight_mean_ = mean(weight_mean),
    err_ = sd(weight_mean) / sqrt(n())
  )

# get number of pellets / delta bw
# merge intake over days with weight data
intake_weight <- intake_over_days %>% 
  left_join(weight_uncertainty %>% 
              select(animal, weight, date), by = c("animal", "ymd" = "date"))
# impute missing data with a rolling average
intake_weight_imputation <- intake_weight %>% 
  group_by(
    animal
  ) %>% 
  mutate(
    rolling_weight = zoo::rollapply(weight, 5, mean, fill = 0, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  filter(rolling_weight != 0) %>% 
  group_by(
    animal,
    session
  ) %>% 
  mutate(
    pellets_per_gr = pellets_per_day / rolling_weight
  ) 
# calculate means per session
intake_weight_session <- intake_weight_imputation %>% 
  group_by(
    session,
    group
  ) %>% 
  summarise(
    pellets_per_gr_mean = mean(pellets_per_gr),
    err = sd(pellets_per_gr) / sqrt(n())
  )
# calculate means per group
intake_weight_group <- intake_weight_session %>% 
  group_by(
    group
  ) %>% 
  summarise(
    pellets_per_gr_mean_group = mean(pellets_per_gr_mean),
    err = sd(pellets_per_gr_mean) / sqrt(n())
  )

# same idea but with light dark periods
# get number of pellets / delta bw
# merge intake over days with weight data
intake_weight_circ <- intake_over_days_circ %>% 
  left_join(weight_uncertainty %>% 
              select(animal, weight, date), by = c("animal", "ymd" = "date"))
# impute missing data with a rolling average, add from the previously calculated
intake_weight_imputation_circ <- intake_weight_circ %>% 
  select(-session) %>% 
  left_join(intake_weight_imputation %>% 
              select(
                animal,
                ymd,
                rolling_weight
              ), by = c("animal", "ymd")) %>% 
  ungroup() %>% 
  filter(rolling_weight != 0) %>% 
  group_by(
    animal,
    session,
    light_dark
  ) %>% 
  mutate(
    pellets_per_gr = pellets_per_day / rolling_weight
  ) 

# calculate means per session
intake_weight_session_circ <- intake_weight_imputation_circ %>% 
  group_by(
    session,
    group,
    light_dark
  ) %>% 
  summarise(
    pellets_per_gr_mean = mean(pellets_per_gr),
    err = sd(pellets_per_gr) / sqrt(n())
  )
# calculate means per group
intake_weight_group_circ <- intake_weight_session_circ %>% 
  group_by(
    group,
    light_dark
  ) %>% 
  summarise(
    pellets_per_gr_mean_group = mean(pellets_per_gr_mean),
    err = sd(pellets_per_gr_mean) / sqrt(n())
  )
```
## lickometer

### preference
```{r}
# add relevant labels: experimental group and protocol
labels <- weight_uncertainty %>% 
  select(animal, group, protocol, experiment_id) %>% 
  distinct(., animal, .keep_all = TRUE) %>% 
  rename(ID = animal)


lickometer_data <- lickometer_uncertainty %>% 
  left_join(
    labels,
    by = c("ID")
  ) %>% 
  # add session instead of dates
  group_by(
    group,
    task_type
  ) %>% 
  mutate(
    session = as.numeric(as.factor(fecha))
  ) %>% 
  ungroup()

# preference for sucrose
spout_pref <- lickometer_data %>% 
  group_by(
    ID,
    group,
    protocol,
    tipo_recompensa,
    session,
    task_type,
    fecha,
    experiment_id
  ) %>% 
  summarise(
    n_licks = n(),
    n_events = max(evento)
  )

# preference for sucrose in FR
spout_pref_fr <- spout_pref %>% 
  filter(task_type == "fr") %>% 
  group_by(
    group,
    protocol,
    tipo_recompensa,
    session,
    fecha
  ) %>% 
  summarise(
    mean_licks = mean(n_licks)
  ) %>% 
  ungroup() %>% 
  group_by(
    group,
    protocol,
    fecha
  ) %>% 
  mutate(
    total_licks = sum(mean_licks),
    relative_licks = (mean_licks / total_licks) * 100
  ) %>% 
  ungroup() %>% 
  group_by(
    group,
    protocol,
    tipo_recompensa
  ) %>% 
  summarise(
    mean_licks_group = mean(relative_licks), 
    err = sd(relative_licks) / sqrt(n())
  )
```

### PR structure
```{r}
pr_sessions <- spout_pref %>% 
  filter(task_type == "pr") %>% 
  group_by(
    group,
    protocol,
    experiment_id
  ) %>% 
  mutate(
    session = as.numeric(as.factor(fecha)),
    valid_session = if_else(protocol == "control" & session > 5, 0, 1)
  ) %>% 
  # consider only sucrose licks and events
  filter(valid_session == 1, tipo_recompensa == "sacarosa") %>% 
  select(-valid_session) %>% 
  # re compute standarized mean after merging both controls group together
  mutate(
    experiment_id = if_else(protocol == "control", "control", experiment_id)
  ) %>% 
  # z-score licks and events
  group_by(
    ID
  ) %>% 
  mutate(
    n_licks_z = scale(n_licks),
    n_events_z = scale(n_events)
  ) %>% 
  ungroup() %>% 
  # obtain group estimates
  group_by(
    protocol,
    group,
    experiment_id,
    session
  ) %>% 
  summarise(
    mean_licks = mean(n_licks_z),
    mean_events = mean(n_events_z),
    err_licks = sd(n_licks_z) / sqrt(n()),
    err_events = sd(n_events_z) / sqrt(n())
  )

# obtain timeout measure via peri-event interval
peri_event_data <- lickometer_data %>% 
  uncumulate()

#saveRDS(peri_event_data %>%  synch_lick_event(., TRUE), "peri_event.rds")
peri_event_data <- readRDS("peri_event.rds") %>% 
  data.table::rbindlist(., fill = TRUE) %>% 
  group_by(ID, n_sesion) %>% 
  mutate(event_id = as.numeric(as.factor(event_id))) %>% 
  ungroup()

peri_event_data_pr <- peri_event_data %>% 
  select(
    experiment_id,
    tiempo_peri_evento,
    task_type,
    protocol,
    fecha,
    ID
  ) %>% 
  filter(
    task_type == "pr"
  ) %>% 
  # merge control groups together
  mutate(
    experiment_id = if_else(protocol == "control", "control", experiment_id)
  ) 

peri_event_data_pr_bin <- peri_event_data_pr %>% 
  group_by(
    ID,
    fecha,
    experiment_id
  ) %>% 
  # create 100 ms bins for peri-event time
  mutate(
    peri_breaks = cut(
        tiempo_peri_evento,
        breaks = seq(-5000, 5000, by = 500),
        labels = FALSE,
        include.lowest = TRUE)) %>% 
  drop_na() %>% 
  ungroup() %>% 
  group_by(
    ID,
    fecha,
    experiment_id,
    peri_breaks
  ) %>% 
  summarise(
    n_licks = n()
  ) %>% 
  ungroup() %>% 
  # get group means
  group_by(
    experiment_id,
    peri_breaks
  ) %>% 
  summarise(
    mean_n_licks = mean(n_licks), 
    err = sd(n_licks) / sqrt(n())
  )

# cum sum data
peri_event_data_pr_bin_cum <- peri_event_data_pr_bin %>% 
  group_by(
    experiment_id
  ) %>% 
  mutate(
    cum_licks = cumsum(mean_n_licks),
    cum_err = cumsum(err)
  )

pr_bin_data <- lickometer_data %>% 
  filter(task_type == "pr", tipo_recompensa == "sacarosa") %>% 
  mutate(
    session = as.numeric(as.factor(fecha)),
    valid_session = if_else(protocol == "control" & session > 5, 0, 1)
  ) %>% 
  # consider only sucrose licks and events
  filter(valid_session == 1, tipo_recompensa == "sacarosa") %>% 
  select(-valid_session) %>% 
  # re compute standarized mean after merging both controls group together
  mutate(
    experiment_id = if_else(protocol == "control", "control", experiment_id)
  ) %>% 
  bin_calculation(., 60, 5) %>% 
  group_by(
    ID,
    session,
    bins,
    experiment_id
  ) %>% 
  summarise(
    n_licks = n()
  ) %>% 
  # group level estimates
  ungroup() %>% 
  group_by(
    session,
    bins,
    experiment_id
  ) %>% 
  summarise(
    m_licks = mean(n_licks)
  ) %>% 
  ungroup() %>% 
  group_by(
    bins,
    experiment_id
  ) %>% 
  summarise(
    mean_licks = mean(m_licks),
    err = sd(m_licks) / sqrt(n())
  ) %>% 
  drop_na()
```

# plots

## intake

### cummulative intake over days
```{r}
# cummulative intake over days
intake_over_days_group %>% 
  ggplot(aes(
    session,
    pellets_cumm_group,
    color = group,
    ymin = pellets_cumm_group - err,
    ymax = pellets_cumm_group + err
  )) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.3)
```
### mean intake per group
```{r}
intake_means %>% 
  ggplot(aes(
    group,
    pellets_mean_group,
    ymin = pellets_mean_group - err,
    ymax = pellets_mean_group + err
  )) +
  geom_col() +
  geom_errorbar(width = 0.3)
```
### pellet intake per bw per session
```{r}
intake_weight_session %>% 
  ggplot(aes(
    session,
    pellets_per_gr_mean,
    color = group,
    ymin = pellets_per_gr_mean - err,
    ymax = pellets_per_gr_mean + err
  )) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.3)
```
### pellet intake per bw per group
```{r}
intake_weight_group %>% 
  ggplot(aes(
    group,
    pellets_per_gr_mean_group,
    ymin = pellets_per_gr_mean_group - err,
    ymax = pellets_per_gr_mean_group + err
  )) +
  geom_col() +
  geom_errorbar(width = 0.3)
```
### pellet intake per bw per group light/dark
```{r}
intake_weight_group_circ %>% 
  ggplot(aes(
    group,
    pellets_per_gr_mean_group,
    ymin = pellets_per_gr_mean_group - err,
    ymax = pellets_per_gr_mean_group + err,
    fill = light_dark
  )) +
  geom_col(position = "dodge") +
  geom_errorbar(width = 0.3, position = position_dodge(width = 0.9))
```

## weight

### delta body weight per session
```{r}
weight_uncertainty_group %>% 
  ggplot(aes(
    session,
    weight_mean,
    ymin = weight_mean - err,
    ymax = weight_mean + err,
    color = group
  )) +
  geom_point() +
  geom_line() + 
  geom_errorbar()
```
### delta body weight per group
```{r}
weight_uncertainty_group_ %>% 
  ggplot(aes(
    group,
    weight_mean_,
    ymin = weight_mean_ - err_,
    ymax = weight_mean_ + err_
  )) +
  geom_col() +
  geom_errorbar(width = 0.3)
```
## inter pellet recovery time
```{r}
inter_pellet_recovery_ind %>% 
  ggplot(aes(
    experiment_id,
    mean_recovery_id,
    group = protocol
  )) +
  geom_col(
    inherit.aes = FALSE,
    aes(
      experiment_id,
      mean_recovery_group,
      fill = protocol
    ),
    data = inter_pellet_recovery_group,
    position = "dodge"
  ) +
  geom_errorbar(
    inherit.aes = FALSE,
    aes(
      experiment_id,
      mean_recovery_group,
      fill = protocol,
      ymin = mean_recovery_group - err,
      ymax = mean_recovery_group + err
    ),
    data = inter_pellet_recovery_group,
    position = position_dodge(0.9),
    width = 0.3
  ) +
  geom_point(position = position_dodge(0.9), color = "gray70")
```

## lickometer plots

### preference for sucrose in fr
```{r}
spout_pref_fr %>% 
  ggplot(aes(
    group,
    mean_licks_group,
    ymin = mean_licks_group - err,
    ymax = mean_licks_group + err,
    fill = tipo_recompensa
  )) +
  geom_col(position = "dodge") +
  geom_errorbar(position = position_dodge(0.9), width = 0.3)
```
### progressive ratio licks over sessions
```{r}
pr_sessions %>% 
  ggplot(aes(
    session,
    mean_licks,
    ymin = mean_licks - err_licks,
    ymax = mean_licks + err_licks,
    color = experiment_id
  )) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.3) +
  facet_wrap(~group, scales = "free_x")
```
### progressive ratio events over sessions
```{r}
pr_sessions %>% 
  ggplot(aes(
    session,
    mean_events,
    ymin = mean_events - err_events,
    ymax = mean_events + err_events,
    color = experiment_id
  )) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.3) +
  facet_wrap(~group, scales = "free_x")
```
### progressive ratio licks in T.O over sessions
```{r}
peri_event_data_pr_bin %>% 
  ggplot(aes(
    peri_breaks,
    mean_n_licks,
    color = experiment_id,
    ymin = mean_n_licks - err,
    ymax = mean_n_licks + err
  )) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(fill = experiment_id)) +
  geom_vline(xintercept = 10.5)
```

### progressive ratio licks in T.O over session cumulative
```{r}
peri_event_data_pr_bin_cum %>% 
  ggplot(aes(
    peri_breaks,
    cum_licks,
    color = experiment_id,
    ymin = cum_licks - cum_err,
    ymax = cum_licks + cum_err
  )) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(fill = experiment_id)) +
  geom_vline(xintercept = 10.5)
```

### licks over time bins
```{r}
pr_bin_data %>% 
  ggplot(aes(
    bins,
    mean_licks,
    ymin = mean_licks - err,
    ymax = mean_licks + err,
    color = experiment_id
  )) +
  geom_line() +
  geom_point() +
  geom_errorbar()
```

